A continuous urbanization index to assess phylogenetic adaptations in urban bird living

### 2018-05-07 ms.Rmd
slowly adding structure to discussion (very slowly)

author: |
  | Corey T. Callaghan^[Corresponding author: c.callaghan@unsw.edu.au] $^1$, Richard E. Major $^1$$^,$$^2$, John H. Wilshire $^1$, John M. Martin $^1$$^,$$^3$, Richard T. Kingsford $^1$, William K. Cornwell $^1$$^,$$^4$
  | $^1$Center for Ecosystem Science, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia; $^2$Australian Museum Research Institute, Sydney, NSW, Australia; $^3$Royal Botanic Gardens and Domain Trust, Sydney, NSW, Australia; $^4$Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia
date: ""
output:
  rmarkdown::pdf_document:
    fig_caption: yes    
    includes:  
      in_header: preamble-latex.tex
fontsize: 12pt
link-citations: no
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/ecology-letters.csl
bibliography: references.bib

@@

Traditionally, the literature which aims to associate avian species' traits with urbanization has been limited, and as such has lead to less generalizable patterns. The majority of studies in this space have focused on discrete classifications of species into adapters, avoiders, and exploiters [e.g., ], or even binary classification based on presence/absence in cities or towns [e.g., ]. We highlight that our analysis used a total of 477 species (Appendix 4), which is greater than many previous studies.

 Our results suggest that urbanization selects against specialist species.

Generalists are ultimately better prepared to thrive, survive, and adapt within urban areas.

We provide the largest study to date using a continuous urbanization metric in order to assess biological and ecological traits which predict a species' adaptability to urban environments.

Global broad-scale empirical datasets are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global, @aronson2016hierarchical, @gutierrez2018citizen], in large part due to the growing use of citizen science data [@silvertown2009new] in combination with rapidly expanding remote sensing technology [@gorelick2017google].

 is rapidly expanding our understanding of urbanization impacts on birds [e.g., ] Understanding specific species' traits which are associated with urban birds allows us to fully understand how to best preserve and maintain avian diversity in cities. We recommend that urban areas should focus on providing habitats which allow granivores, insectivores, and others to be found more readily in cities.

Our study looked over a broad spatial and temporal scale to identify a novel approach to quantifying a continuous urbanization index for birds. We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global], but migratory status was considered as a predictor. We highlight that future studies, especially within migrant-dominated landscapes (i.e., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [CITATIONS]. Further, future studies should look to expand this analysis outside of the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2017non]). Future studies should aim to (1) broaden the spatial scale, (2) investigate the temporal changes in a species' urbanization indices [@la2014role, @lepczyk2017global], and (3) investigate the role that city size plays on selected traits and phylogenetic diversity [@sol2017urbanisation, @batary2017non].

Ultimately, we highlight the applicability of broad-scale empirical data to assess species' traits which predict presence and adaptability in urban environments. Our results suggest that generalist species are better prepared to forage and survive in the ecologically novel aspects of urban environments [CITATIONS]. Australia is a relatively young country and fully understanding how birds respond to urban development is critical to inform future management. Urban avian biodiversity could be increased by improving the suitability of foraging for granivores and insectivores, especially important given the recent noted collapse of insects in other parts of the world [CITATION]. Furthermore, future studies should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved modelling performance.

>>>>

author: |
  | Corey T. Callaghan^[Corresponding author: c.callaghan@unsw.edu.au] $^1$, Richard E. Major $^1$$^,$$^2$, John H. Wilshire $^1$, John M. Martin $^1$$^,$$^3$, Richard T. Kingsford $^1$, William K. Cornwell $^1$$^,$$^4$
  | $^1$Centre for Ecosystem Science, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia; $^2$Australian Museum Research Institute, Sydney, NSW, Australia; $^3$Royal Botanic Gardens and Domain Trust, Sydney, NSW, Australia; $^4$Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia
date: ""
output:
  rmarkdown::pdf_document:
    fig_caption: yes    
    includes:  
      in_header: preamble-latex.tex
fontsize: 12pt
link-citations: no
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/ecology-letters.csl
bibliography: references.bib

@@

**A PARAGRAPH ON SPECIFIC TRAITS AND CITATIONS WHICH CONFIRM OR DENY THEM**

 Fully understanding specific species' traits which are associated with urban birds allows us to fully understand how to best preserve and maintain avian biodiversity in urban areas [@evans2011makes]. However, previous studies which have assessed this have generally focused on discrete classifications of species based on their responses to urbanization, limiting the applicability of results found. We further highlight that our analysis used a total of 477 species (Appendix 4), greater than many previous studies, ultimately allowing for more generalizable results. The aforementioned differences in our study compared with others is likely a result of our increased sample size, as well as our use of a continuous measure of urbanization.

Our results suggest that urbanization selects against specialist species, granivores, and insectivores. As such, future urban development should focus on maintaining a high diversity of habitat (CITATIONS) both for foraging (CITATION) and nesting (CITATIONS). Further, a focus should be on providing habitat for insects (CITATION), which will in turn increase the presence of insectivores in urban environments.

Global broad-scale empirical datasets are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global, @aronson2016hierarchical, @gutierrez2018citizen], in large part due to the growing use of citizen science data [@silvertown2009new] in combination with advancing remote sensing technology [@gorelick2017google]. We used a novel approach, over a broad spatial and temporal scale, to quantify an urbanization index in birds. We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global], but migratory status was considered as a predictor. We highlight that future studies, especially within migrant-dominated landscapes (i.e., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full, @marra2015call]. Further, future studies should look to expand this analysis outside of the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2017non]).

Ultimately, we provide the largest study to date using a continuous urbanization metric in order to assess biological and ecological traits which predict a species' adaptability to urban environments. In doing so, we highlight the applicability of broad-scale empirical data to assess species' traits which predict presence and adaptability in urban environments. Our results suggest that generalist species are better prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns]. Australia is a relatively young country and fully understanding how birds respond to urban development is critical to inform future management. Urban avian biodiversity could be increased by improving the suitability of foraging for granivores and insectivores, especially important given the recent noted collapse of insects in other parts of the world [CITATION]. Furthermore, future studies should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved modelling performance.
